Multi-wavelength Radio Continuum Emission Studies of Dust-free Red Giants

### 2013-04-27 draft_07/draft.tex
27/04 More co-I edits

Multi-wavelength centimeter continuum observations of non-dusty, non-pulsating K spectral-type red giants directly sample their chromospheres and wind acceleration zones where most of the energy that drives their mass loss is deposited. Such stars are feeble emitters at these wavelengths however, and previous observations have provided only a small number of modest signal-to-noise measurements slowly accumulated over three decades. We present multi-wavelength Karl G. Jansky Very Large Array (VLA) thermal continuum observations of the wind acceleration zones of two dust-free red giants, Arcturus ($\alpha$ Boo: K2 III) and Aldebaran ($\alpha$ Tau: K5 III). Importantly, most of our observations of each star were carried out over just a few days, so that we obtained an essentially consistent snapshot of the different stellar layers sampled at different wavelengths, independent of any long-term variability. We report the first detections at several wavelengths for each star including a detection at 10 cm (3.0 GHz: S-band) for both stars and a 20 cm (1.5 GHz: L-band) detection for $\alpha$ Boo. This is the first time single (non-binary) luminosity class III red giants have been detected at these continuum wavelengths. We find that these flux density measurements sample the outer layers of $\alpha$ Boo's atmosphere where its wind velocity is approaching (or possibly has reached) its terminal value and the ionization balance is becoming \textit{frozen-in}. We present spectral energy distributions for both stars and compare these with previous measurements and published semi-empirical models based on ultraviolet data. The marked deviations between our data and existing semi-empirical models highlight the need for new atmospheric models to be developed. Spectral indices are used to discuss the possible properties of the stellar atmospheres and we find evidence for a rapidly cooling wind in the case of $\alpha$ Boo. We develop a simple analytical atmospheric model for $\alpha$ Boo which includes a rapidly cooling constant velocity wind, and the model is found to be in good agreement with our new long wavelength flux density values.

@@

Although studies of wind-scattered UV and optical line profiles have provided clues to the mass loss rates and radial distribution of the mean and turbulent velocity fields, the thermal structure remains poorly constrained. In the UV, the source function of electron collisionally excited emission lines is sensitive to electron temperature (i.e., $S_{\nu} \propto e^{-h\nu /kT}$) and so a localized hot plasma component in a dynamic atmosphere can completely dominate the temporally and spatially averaged emission and not hence reflect the global mean value. At radio wavelengths however, the source function is thermal and is just the Rayleigh-Jeans tail of the Planck function, which is linear in electron temperature (i.e., $S_{\nu} = {2kT\nu ^2 /c^2}$) and so should give a more appropriate estimate of the mean thermal wind structure. It is this value that controls the atomic level populations and ionization of the mean plasma that feed into UV spectroscopic analyses. This value is also needed to quantify the implied thermal heating supplied to the  wind by the unknown driving source/sources, allowing constraints on potential mass loss mechanisms to be derived.

In the cm-radio regime the radio opacity strongly increases with wavelength (i.e., $ \kappa _{\lambda} \propto \lambda ^{2.1}$) and so the longer wavelengths sample the extended layers of a stars atmosphere, thus	 providing us with spatial information about the star's mass outflow region. The NRAO\footnote{The National Radio Astronomy Observatory is a facility of the National Science Foundation operated under cooperative agreement by Associated Universities, Inc.} Karl G. Jansky Very Large Array (VLA) is sensitive to over three orders of magnitude in continuum optical depth ($\tau _{20 \ \rm{cm}}/\tau _{0.7 \ \rm{cm}} \approx 10^3$) and provides an area-averaged sweep through the wind acceleration zone of evolved late type stars. The thermodynamic properties in this spatial region control the ionization in the far wind because the ionization balance, which also controls the cooling rates, becomes \textit{frozen-in} at large radii due to advection. Furthermore, it is these outer extended regions of the star's atmosphere that contribute to the commonly seen P Cygni line profiles in the UV. In these profiles the line-of-sight absorption caused by the star's wind is superimposed on the blueshifted scattered emission. Thus, centimeter radio continuum observations can provide a test of models based on these UV profiles. In this paper we directly compare our new VLA observations with atmospheric models derived from UV analysis.

@@

\begin{deluxetable}{ccc}
\tabletypesize{\scriptsize}
\tablecaption{Properties of red giant sample.}
\tablehead{	\colhead{}		       				& 
			\colhead{$\alpha$ Boo}				&
			\colhead{$\alpha$ Tau}				}
\startdata
Spectral Type 				& K2 III  & K5 III  \\
HD Number 				& 124897  & 29139  \\
Mass (M_{\odot})	& $0.8 \pm 0.2$  & $1.3 \pm 0.3$ \\
Effective Temperature (K)	& $4294 \pm 30$  & $3970 \pm 49$ \\
Angular Diameter (mas)		& $21.0 \pm 0.2$ & $20.2 \pm 0.3$ \\
Distance (pc)	& $11.3 \pm 0.1$ & $20.4 \pm 0.3$\\
Radius ($R_{\odot}$)	& $25.4 \pm 0.3$  & $44.4 \pm 1.0$ \\
Photospheric Escape Velocity & 110 km s$^{-1}$ & 106 km s$^{-1}$ \\
Wind Terminal Velocity & $\sim$40 km s$^{-1}$ & $\sim$30 km s$^{-1}$ \\
Mass Loss Rate ($M_{\odot}$ yr$^{-1}$)		& 2$\times$10$^{-10}$ & 1.6$\times$10$^{-11}$ \\
Wind Temperature (K)		& $\sim$10,000  & $< 10,000$  \\
Semi-empirical Model	& \cite{1985pssl.proc..351D} & \cite{1999MNRAS.302...37M}
\enddata
\tablecomments{Masses are from \cite{2010A&A...509A..77K} and \cite{2012A&A...547A.108L}. Effective temperatures and photospheric angular diameters are from \cite{1993AA...270..315D}. Distances are from \cite{2007A&A...474..653V}. Mass loss rates are from \cite{1985pssl.proc..351D} and \cite{1998ApJ...503..396R}.}

@@

\begin{deluxetable*}{ccccccccc}
\tabletypesize{\scriptsize}
%\tablecolumns{6} 
%\tablewidth{0pt} 
\tablecaption{VLA Observations}
\tablehead{\colhead{Star}			            &
\colhead{Date}
				&
          	\colhead{Band}	      				&
          	\colhead{Frequency\tablenotemark{a}}&
	\colhead{Wavelength}			         	&
           	\colhead{Time on Star}              &
           	\colhead{Bandwidth}            		&
           	\colhead{Number of}            		&
	\colhead{Phase}		\\
	\colhead{}									&
    \colhead{}		                			& 
    \colhead{}		                			& 
	\colhead{(GHz)}                        		& 
	\colhead{(cm)}                         		& 
	\colhead{(hr)}                    			&
	\colhead{(GHz)}                        		&
	\colhead{Antennae\tablenotemark{b}}         &
	\colhead{Calibrator}		}
\startdata
$\alpha$ Boo	& 2011 Feb 22 & Q	& 43.3 & 0.7		& 0.3 	&  0.256	&22& J1357+1919  \\
$\alpha$ Boo 	& 2011 Feb 22 & Ka	& 33.6 & 0.9		& 0.2 	&  0.256 	&23&J1357+1919  \\
$\alpha$ Boo 	& 2011 Feb 22 & K	& 22.5 & 1.3		& 0.4	&  0.256 	&24&J1357+1919  \\
$\alpha$ Boo 	& 2011 Feb 11 & X	& 8.5  & 3.5		& 0.3 	&  0.256 	&18&J1415+1320  \\
$\alpha$ Boo	& 2011 Feb 11 & C	& 5.0  & 6.0 		& 0.5	& 0.256 	&21& J1415+1320 \\
$\alpha$ Boo	& 2011 Feb 13 & S	& 3.1  & 9.5 		& 1.8 	& 0.256 	&12& J1415+1320 \\
$\alpha$ Boo 	& 2012 Jul 19 & S	& 3.0  & 10.0 		& 0.7 	& 2.0 		&23& J1415+1320 \\
$\alpha$ Boo	& 2012 Jul 20 & L	& 1.5  & 20.0		& 1.6 	& 1.0 		&23& J1415+1320 \\
$\alpha$ Tau	& 2011 Feb 11 & Q	& 43.3 & 0.7 		& 0.3 	& 0.256 	&22&  J0431+1731\\
$\alpha$ Tau	& 2011 Feb 11 & Ka	& 33.6 & 0.9 		& 0.2 	& 0.256 	&19&  J0449+1121\\
$\alpha$ Tau	& 2011 Feb 11 & K	& 22.5 & 1.3 		& 0.4 	& 0.256 	&21&  J0449+1121\\
$\alpha$ Tau	& 2011 Feb 13 & X	&  8.5 & 3.5 		& 0.5	& 0.256 	&25&  J0449+1121\\
$\alpha$ Tau	& 2011 Feb 13 & C	&  5.0 & 6.0 		& 1.2	& 0.256 	&21&  J0449+1121\\
$\alpha$ Tau	& 2011 Feb 12 & S	&  3.1 & 9.5 		& 1.8 	& 0.256 	&11&  J0431+2037
\enddata
\tablenotetext{a}{Central frequency of selected bandpass.}
\tablenotetext{b}{Number of available antennae remaining after flagging.}
\label{tab:tab2}
\end{deluxetable*}

@@

Both $\alpha$ Boo and $\alpha$ Tau were slightly offset from the phase-center by $\sim$5 synthesized beam widths in order to avoid possible errors at phase-center. All scheduling blocks (SBs) were kept to $\le$\,2.5 hours of duration. For the high frequency observations (i.e., Q, Ka, and K-bands) we used the \textit{fast switching} technique which consists of rapidly alternating observations of the target source and a nearby unresolved phase calibrator. The total cycle times for the Q, Ka, and K-band observations were 160, 230, and 290 s, respectively. For both target sources these high frequency observations were combined into a single 2 hour observing track and commenced with X-band reference pointing with solutions being applied on-line. After X-band pointing the target source was observed at Q-band to insure the best pointing solutions were used. The tracks at lower frequencies were composed of repeatedly interleaving observations of the target source and a nearby phase calibrator but had longer cycle times. The primary calibration sources 3C286 and 3C138 were observed at the end of all tracks and were used to measure the complex bandpass and set the absolute flux for $\alpha$ Boo and $\alpha$ Tau, respectively.

@@

\begin{deluxetable*}{Cccccccc}
\tabletypesize{\scriptsize}
%\tablecolumns{6} 
%\tablewidth{0pt} 
\tablecaption{VLA Flux Densities of $\alpha$ Boo and $\alpha$ Tau}
\tablehead{\colhead{}		            &
			\colhead{Band}		            &
			\colhead{Frequency\tablenotemark{a}}&
			\colhead{Peak $F_{\nu}$}					&
          	\colhead{Integrated $F_{\nu}$} 			&
          	\colhead{\textit{Imfit} Integrated}	&
			\colhead{Image rms}					&
           	\colhead{\textit{Imfit} Fitting}  	\\
	\colhead{}									&
	\colhead{}									&
	\colhead{(GHz)}								&
    \colhead{(mJy)}		        	& 
    \colhead{(mJy)}		                		& 
	\colhead{$F_{\nu}$ (\rm{mJy})}                        		& 
	\colhead{(Jy beam^{-1})}         					&
	\colhead{Error (mJy)}		}
\startdata
$\alpha$ Boo (K2 III) &Q  &43.28& 5.941 & 6.093 & 6.420 & 0.301 &  0.261\\
&Ka &33.56& 4.159 & 4.319 & 4.489 & 0.083 & 0.090 \\
&K  &22.46& 1.827 & 1.784 & 1.809 & 0.043 & 0.050 \\
&X  &8.46 & 0.510 & 0.514 & 0.532 & 0.030 & 0.019 \\
&C  &4.90 & 0.214 & 0.144 & 0.159 & 0.035 & 0.009 \\
&S  &3.15 & 0.148 & 0.135 & -     & 0.028 & -     \\
&S  &2.87 & 0.127 & 0.118 & 0.116 & 0.012 & 0.016\\
&L  &1.63 & 0.067 & 0.068 & -     & 0.013 & -    \\
\hline
\rule{0pt}{3ex}  $\alpha$ Tau (K5 III)&Q  &43.28 & 3.672& 3.734 & 4.082 &  0.259& 0.183	\\
&Ka &33.56 & 2.188& 1.962 & 2.125 &  0.091& 0.070 \\
&K  &22.46 & 1.864& 1.881 & 2.069 &  0.042& 0.083 \\
&X  &8.46  & 0.296& 0.287 & 0.281 &  0.014& 0.015 \\
&C  &4.96  & 0.147& 0.167 & 0.176 &  0.010& 0.010 \\
&S  &3.15  & 0.062& 0.043 & - &  0.017 & -
\enddata
\tablenotetext{a}{Frequency of the final image produced using the multi-frequency synthesis imaging mode within CASA's \textit{clean} task.}
\label{tab:tab3}
\end{deluxetable*}

@@

\subsection{$\alpha$ Boo Radio Maps} \label{results1} 
High S/N detections ($>19\sigma$) of $\alpha$ Boo were made at 22.5, 33.6, and 43.3 GHz. Some residuals of the dirty beam remained in the CLEANed maps due to the paucity of uv-coverage in these short high frequency observations. The lower-frequency maps were contaminated by the emission of a strong radio source located $186\arcsec$ north-west of $\alpha$ Boo. This source was found to have a flux density of 25 mJy at 4.9 GHz by \cite{1986AJ.....91..602D}, in close agreement with our measurement of 23.2 mJy at the same frequency. We find the source to have a spectral index $\alpha$ ($F_{\nu} \propto \nu ^{\alpha}$), equal to -1.4 and reaches 80.3 mJy at 1.6 GHz. For unknown reasons we do not detect $\alpha$ Boo in the higher frequency sub-band at 5.0 GHz (C-band) so the values given in Table \ref{tab:tab3} are taken from the lower frequency sub-band only. We obtain good detections ($>5\sigma$) of the star for both epochs at $\sim$3 GHz (S-band) and the peak flux densities agree within their uncertainties. We can therefore safely assume that the 1.5 GHz (L-band) flux density has not changed significantly over that period either, and so can safely be included in any analysis. The map at L-band was highly contaminated by the sidelobes of the strong source north-west of $\alpha$ Boo but the star is still detected at the $5\sigma$ level. There is a slight positional offset of $1\arcsec$ between the position of the peak flux density at 1.5 and at 3.0 GHz for the 2012 data, which were taken within 1 day of each other. However, the position uncertainties, due to noise and to the atmospheric differences between the directions of the phase reference source and the target, are at least $1\arcsec$, and so we feel that it is highly likely that both detections are of $\alpha$ Boo.

@@

@@

\begin{figure*}
\centering
\includegraphics[trim = 0mm 0mm 0mm 20mm, clip,scale=0.65, angle=90]{fig1.ps}
\\
\caption{Spectral energy distribution of $\alpha$ Boo for 1 GHz $\leq \nu \leq$ 1 THz. Our new multi-frequency VLA observations which were mainly acquired over a few days in February 2011 are the blue circles and are in disagreement with the existing chromospheric and wind models of \cite{1985pssl.proc..351D}. The overlap between the two models are represented by the green shaded area. The red triangles are previous observations which were acquired sporadically over the last three decades with the `old' VLA, IRAM and BIMA. The black dashed line is the expected radio emission from the Drake model which undergoes rapid wind cooling beyond $\sim$2.3 $R_{\star}$ (see Section \ref{disc:disc3} and \ref{disc:disc4}).}
\label{fig:fig1}
\centering
\includegraphics[trim = 0mm 0mm 0mm 20mm, clip,scale=0.65, angle=90]{fig2.ps}
\\
\caption{Spectral energy distribution of $\alpha$ Tau for 1 GHz $\leq \nu \leq$ 1 THz. Our new multi-frequency VLA observations of $\alpha$ Tau (blue circles) were acquired in just two days in February 2011. The red triangles are the previous radio observations of the star which were acquired over many years. The green line is the expected radio emission from the existing hybrid chromosphere and wind model while the dashed  green line is the expected radio emission from a theoretical Alfv\'en wave driven model atmosphere.}
\label{fig:fig2}
\end{figure*}

>>>>

Multi-wavelength centimeter continuum observations of non-dusty, non-pulsating K spectral-type red giants directly sample their chromospheres and wind acceleration zones where most of the energy that drives their mass loss is deposited. Such stars are feeble emitters at these wavelengths however, and previous observations have provided only a small number of modest signal-to-noise measurements slowly accumulated over three decades. We present multi-wavelength Karl G. Jansky Very Large Array (VLA) thermal continuum observations of the wind acceleration zones of two dust-free red giants, Arcturus ($\alpha$ Boo: K2 III) and Aldebaran ($\alpha$ Tau: K5 III). Importantly, most of our observations of each star were carried out over just a few days, so that we obtained an essentially consistent snapshot of the different stellar layers sampled at different wavelengths, independent of any long-term variability. We report the first detections at several wavelengths for each star including a detection at 10 cm (3.0 GHz: S-band) for both stars and a 20 cm (1.5 GHz: L-band) detection for $\alpha$ Boo. This is the first time single (non-binary) luminosity class III red giants have been detected at these continuum wavelengths. We sample the outer layers of $\alpha$ Boo's atmosphere where its wind velocity is approaching (or possibly has reached) its terminal value and the ionization balance is becoming \textit{frozen-in}. For $\alpha$ Tau however, our long-wavelength data is still sampling its inner atmosphere where the wind is still accelerating. We present spectral energy distributions for both stars and compare these with previous measurements and published semi-empirical models based on ultraviolet data. The marked deviations between our data and existing semi-empirical models highlight the need for new atmospheric models to be developed. Spectral indices are used to discuss the possible properties of the stellar atmospheres and we find evidence for a rapidly cooling wind in the case of $\alpha$ Boo. We develop a simple analytical atmospheric model for $\alpha$ Boo which includes a rapidly cooling constant velocity wind, and the model is found to be in good agreement with our new long wavelength flux density values.

@@

Although studies of wind-scattered UV and optical line profiles have provided clues to the mass loss rates and radial distribution of the mean and turbulent velocity fields, the thermal structure remains poorly constrained. In the UV, the source function, $S_{\nu}$, of electron collisionally excited emission lines is sensitive to electron temperature, $T_{\rm{e}}$, (i.e., $S_{\nu} \propto e^{-h\nu /kT_{\rm{e}}}$). Therefore, a localized hot plasma component in a dynamic atmosphere can completely dominate the temporally and spatially averaged emission and hence not reflect the global mean radial electron temperature value. At radio wavelengths however, the source function is thermal and is just the Rayleigh-Jeans tail of the Planck function, which is linear in electron temperature (i.e., $S_{\nu} = {2kT_{\rm{e}}\nu ^2 /c^2}$). This should give a more appropriate estimate of the mean radial electron temperature. It is this value that controls the atomic level populations and ionization of the mean plasma that feed into UV spectroscopic analyses. This value is also needed to quantify the implied thermal heating supplied to the  wind by the unknown driving source/sources, allowing constraints on potential mass loss mechanisms to be derived.

In the cm-radio regime the radio opacity, $ \kappa _{\lambda}$, strongly increases with wavelength (i.e., $ \kappa _{\lambda} \propto \lambda ^{2.1}$) and so the longer wavelengths sample the extended layers of a stars atmosphere, thus	 providing us with spatial information about the star's mass outflow region. The NRAO\footnote{The National Radio Astronomy Observatory is a facility of the National Science Foundation operated under cooperative agreement by Associated Universities, Inc.} Karl G. Jansky Very Large Array (VLA) is sensitive to over three orders of magnitude in continuum optical depth, $\tau _{\lambda}$, ($\tau _{20 \ \rm{cm}}/\tau _{0.7 \ \rm{cm}} \approx 10^3$) and provides an area-averaged sweep through the wind acceleration zone of evolved late type stars. The thermodynamic properties in this spatial region control the ionization in the far wind because the ionization balance, which also controls the cooling rates, becomes \textit{frozen-in} at large radii due to advection. Furthermore, it is these outer extended regions of the star's atmosphere that contribute to the commonly seen P Cygni line profiles in the UV. In these profiles the line-of-sight absorption caused by the star's wind is superimposed on the blueshifted scattered emission. Thus, centimeter radio continuum observations can provide a test of models based on these UV profiles. In this paper we directly compare our new VLA observations with atmospheric models derived from UV analysis.

@@

\begin{deluxetable}{lcc}
\tabletypesize{\scriptsize}
\tablecaption{Properties of red giant sample.}
\tablehead{	\colhead{}		       				& 
			\colhead{$\alpha$ Boo}				&
			\colhead{$\alpha$ Tau}				}
\startdata
Spectral Type$\dotfill$	& K2 III  & K5 III  \\
HD Number$\dotfill$				& 124897  & 29139  \\
Mass ($M_{\odot}$)$\dotfill$	& $0.8 \pm 0.2$  & $1.3 \pm 0.3$ \\
Effective Temperature (K)$\dotfill$	& $4294 \pm 30$  & $3970 \pm 49$ \\
Angular Diameter (mas)$\dotfill$		& $21.0 \pm 0.2$ & $20.2 \pm 0.3$ \\
Distance (pc)$\dotfill$	& $11.3 \pm 0.1$ & $20.4 \pm 0.3$\\
Radius ($R_{\odot}$)$\dotfill$	& $25.4 \pm 0.3$  & $44.4 \pm 1.0$ \\
Photospheric Escape Velocity$\dotfill$ & 110 km s$^{-1}$ & 106 km s$^{-1}$ \\
Wind Terminal Velocity$\dotfill$ & $\sim$40 km s$^{-1}$ & $\sim$30 km s$^{-1}$ \\
Mass Loss Rate ($M_{\odot}$ yr$^{-1}$)$\dotfill$	& 2$\times$10$^{-10}$ & 1.6$\times$10$^{-11}$ \\
Wind Temperature (K)$\dotfill$		& $\sim$10,000  & $< 10,000$  \\
Semi-empirical Model$\dotfill$	& \cite{1985pssl.proc..351D} & \cite{1999MNRAS.302...37M}
\enddata
\tablecomments{Masses are from \cite{2010A&A...509A..77K} and \cite{2012A&A...547A.108L}. Effective temperatures and photospheric angular diameters are from \cite{1993AA...270..315D}. Distances are from \cite{2007A&A...474..653V}. Mass loss rates are from \cite{1985pssl.proc..351D} and \cite{1998ApJ...503..396R}.}

@@

\begin{deluxetable*}{lccccccccc}
\tabletypesize{\scriptsize}
%\tablecolumns{6} 
%\tablewidth{0pt} 
\tablecaption{VLA Observations}
\tablehead{\colhead{Star}			            &
\colhead{Date}
				&
          	\colhead{Band}	      				&
          	\colhead{Frequency\tablenotemark{a}}&
	\colhead{Wavelength}			         	&
           	\colhead{Time on Star}            &
           	\colhead{Restoring Beam}              &
           	\colhead{Bandwidth}            		&
           	\colhead{Number of}            		&
	\colhead{Phase}		\\
	\colhead{}									&
    \colhead{}		                			& 
    \colhead{}		                			& 
	\colhead{(GHz)}                        		& 
	\colhead{(cm)}                         		& 
	\colhead{(hr)}                    			&
	\colhead{($\arcsec \times \arcsec$ )}                    			&
	\colhead{(GHz)}                        		&
	\colhead{Antennae\tablenotemark{b}}         &
	\colhead{Calibrator}		}
\startdata
$\alpha$ Boo 	& 2011 Feb 22 & Q	& 43.3 & 0.7		& 0.3 	&0.19 $\times$ 0.15& 0.256	&22& J1357+1919  \\
				& 2011 Feb 22 & Ka	& 33.6 & 0.9		& 0.2 	&0.25 $\times$ 0.20& 0.256 	&23&J1357+1919  \\
				& 2011 Feb 22 & K	& 22.5 & 1.3		& 0.4	&0.35 $\times$ 0.28& 0.256 	&24&J1357+1919  \\
				& 2011 Feb 11 & X	& 8.5  & 3.5		& 0.3 	&1.14 $\times$ 0.70& 0.256 	&18&J1415+1320  \\
				& 2011 Feb 11 & C	& 5.0  & 6.0 		& 0.5	&2.02 $\times$ 1.30& 0.256 	&21& J1415+1320 \\
				& 2011 Feb 13 & S	& 3.1  & 9.5 		& 1.8 	&2.57 $\times$ 2.08& 0.256 	&12& J1415+1320 \\
				& 2012 Jul 19 & S	& 3.0  & 10.0 		& 0.7 	&2.82 $\times$ 2.30& 2.0		&23& J1415+1320 \\
				& 2012 Jul 20 & L	& 1.5  & 20.0		& 1.6 	&4.46 $\times$ 3.94& 1.0		&23& J1415+1320 \\
\hline
\rule{0pt}{3ex}  $\alpha$ Tau	& 2011 Feb 11 & Q	& 43.3 & 0.7 		& 0.3 	&0.18 $\times$0.16& 0.256 	&22&  J0431+1731\\
				& 2011 Feb 11 & Ka	& 33.6 & 0.9 		& 0.2 	&0.22 $\times$ 0.20& 0.256 	&19&  J0449+1121\\
				& 2011 Feb 11 & K	& 22.5 & 1.3 		& 0.4 	&0.35 $\times$ 0.31& 0.256 	&21&  J0449+1121\\
				& 2011 Feb 13 & X	&  8.5 & 3.5 		& 0.5	&0.85 $\times$ 0.78& 0.256 	&25&  J0449+1121\\
				& 2011 Feb 13 & C	&  5.0 & 6.0 		& 1.2	&1.48 $\times$ 1.32& 0.256 	&21&  J0449+1121\\
				& 2011 Feb 12 & S	&  3.1 & 9.5 		& 1.8 	&2.74 $\times$ 2.02& 0.256 	&11&  J0431+2037
\enddata
\tablenotetext{a}{Central frequency of selected bandpass.}
\tablenotetext{b}{Number of available antennae remaining after flagging.}
\label{tab:tab2}
\end{deluxetable*}

@@

Both $\alpha$ Boo and $\alpha$ Tau were slightly offset from the phase-center by $\sim$5 synthesized beam widths in order to avoid possible errors at phase-center. All scheduling blocks (SBs) were kept to $\le$\,2.5 hours of duration. For the high frequency observations (i.e., Q, Ka, and K-bands) we used the \textit{fast switching} technique which consists of rapidly alternating observations of the target source and a nearby unresolved phase calibrator. The total cycle times for the Q, Ka, and K-band observations were 160, 230, and 290 s, respectively. For both target sources these high frequency observations were combined into a single 2 hour observing track and commenced with X-band reference pointing with solutions being applied on-line. After X-band pointing the target source was observed at Q-band to ensure the best pointing solutions were used. The tracks at lower frequencies were composed of repeatedly interleaved observations of the target source and a nearby phase calibrator but had longer cycle times. The primary calibration sources 3C286 and 3C138 were observed at the end of all tracks and were used to measure the complex bandpass and set the absolute flux for $\alpha$ Boo and $\alpha$ Tau, respectively.

@@

\begin{deluxetable*}{cccccccc}
\tabletypesize{\scriptsize}
%\tablecolumns{6} 
%\tablewidth{0pt} 
\tablecaption{VLA Flux Densities of $\alpha$ Boo and $\alpha$ Tau}
\tablehead{\colhead{Star}		            &
			\colhead{Band}		            &
			\colhead{Frequency\tablenotemark{a}}&
			\colhead{Peak $F_{\nu}$}					&
          	\colhead{Integrated $F_{\nu}$} 			&
          	\colhead{\textit{Imfit} Integrated}	&
			\colhead{Image rms}					&
           	\colhead{\textit{Imfit} Fitting Error}  	\\
	\colhead{}									&
	\colhead{}									&
	\colhead{(GHz)}								&
    \colhead{(mJy beam$^{-1}$)}		        	& 
    \colhead{(mJy beam$^{-1}$)}		                		& 
	\colhead{$F_{\nu}$ (\rm{mJy beam$^{-1}$})}                        		& 
	\colhead{(mJy beam$^{-1}$)}         					&
	\colhead{(mJy beam$^{-1}$)}		}
\startdata
$\alpha$ Boo  &Q  &43.28& 5.94 & 6.09 & 6.42 & 0.30 &  0.26\\
&Ka &33.56& 4.16 & 4.32 & 4.49 & 0.08 & 0.09 \\
&K  &22.46& 1.83 & 1.78 & 1.81 & 0.04 & 0.05 \\
&X  &8.46 & 0.51 & 0.51 & 0.53 & 0.03 & 0.02 \\
&C  &4.90 & 0.21 & 0.14 & 0.16 & 0.04 & 0.01 \\
&S  &3.15 & 0.15 & 0.14 & -     & 0.03 & -     \\
&S  &2.87 & 0.13 & 0.12 & 0.12 & 0.01 & 0.02\\
&L  &1.63 & 0.07 & 0.07 & -     & 0.01 & -    \\
\hline
\rule{0pt}{3ex}  $\alpha$ Tau &Q  &43.28 & 3.67& 3.73 & 4.08 &  0.26& 0.18	\\
&Ka &33.56 & 2.19& 1.96 & 2.13 &  0.09& 0.07 \\
&K  &22.46 & 1.86& 1.88 & 2.07 &  0.04& 0.08 \\
&X  &8.46  & 0.30& 0.29 & 0.28 &  0.01& 0.02 \\
&C  &4.96  & 0.15& 0.17 & 0.18 &  0.01& 0.01 \\
&S  &3.15  & 0.06& 0.04 & - &  0.02 & -
\enddata
\tablenotetext{a}{Frequency of the final image produced using the multi-frequency synthesis imaging mode within CASA's \textit{clean} task.}
\label{tab:tab3}
\end{deluxetable*}

@@

\subsection{$\alpha$ Boo Radio Maps} \label{results1} 
High S/N detections ($>19\sigma$) of $\alpha$ Boo were made at 22.5, 33.6, and 43.3 GHz. Some residuals of the dirty beam remained in the CLEANed maps due to the paucity of uv-coverage in these short high frequency observations. At the lower frequencies, it was necessary to image confusing sources, notably a strong radio source located $186\arcsec$ north-west of $\alpha$ Boo. This source was reported by \cite{1986AJ.....91..602D} and their flux density of 25 mJy at 4.9 GHz is in close agreement with our measurement of 23.2 mJy at the same frequency. We find the source to have a spectral index $\alpha$ ($F_{\nu} \propto \nu ^{\alpha}$), equal to -1.4; it reaches 80.3 mJy at 1.6 GHz.

We detected $\alpha$ Boo at 6$\sigma$ in the lower frequency sub-band of C-band, at 4.9 GHz. The noise was slightly higher and the images were poorer quality in the C-band higher frequency sub-band, with artifacts exceeding $\pm$200 $\mu$Jy, and we cannot report a detection in this sub-band, so values given in Table \ref{tab:tab3} are taken from the lower frequency sub-band only. We obtain good detections ($>5\sigma$) of the star for both epochs at $\sim$3 GHz (S-band) and the peak flux densities agree within their uncertainties. We can therefore safely assume that the 1.5 GHz (L-band) flux density has not changed significantly over that period either, and so can safely be included in any analysis. The map at L-band was highly contaminated by the sidelobes of the strong source north-west of $\alpha$ Boo but the star is still detected at the $5\sigma$ level. There is a slight positional offset of $1\arcsec$ between the position of the peak flux density at 1.5 and at 3.0 GHz for the 2012 data, which were taken within 1 day of each other. However, the position uncertainties, due to noise and to the atmospheric differences between the directions of the phase reference source and the target, are at least $1\arcsec$, and so we feel that it is highly likely that both detections are of $\alpha$ Boo.

@@

\begin{figure}
\includegraphics[trim = 0mm 0mm 0mm 0mm, clip,scale=0.36,angle=90]{fig0.ps}
\caption{Existing atmospheric model for $\alpha$ Boo \cite[`model A']{1985pssl.proc..351D} along with the same model which undergoes rapid wind cooling beyond $\sim$2.3 $R_{\star}$ (see Section \ref{disc:disc4}). }
\label{fig:fig0}
\end{figure}

\begin{figure*}
\centering
\includegraphics[trim = 0mm 0mm 0mm 20mm, clip,scale=0.65, angle=90]{fig1.ps}
\\
\caption{Spectral energy distribution of $\alpha$ Boo for 1 GHz $\leq \nu \leq$ 1 THz. Our new multi-frequency VLA observations which were mainly acquired over a few days in February 2011 are the blue circles and disagree with the existing chromospheric and wind models of \cite{1985pssl.proc..351D}. The overlap between the two models is represented by the green shaded area. The red triangles are previous observations which were acquired sporadically over the last three decades with the `old' VLA, IRAM and BIMA. The black dashed line is the expected radio emission from the Drake model which undergoes rapid wind cooling beyond $\sim$2.3 $R_{\star}$ (see Section \ref{disc:disc3} and \ref{disc:disc4}).}
\label{fig:fig1}
\centering
\includegraphics[trim = 0mm 0mm 0mm 20mm, clip,scale=0.65, angle=90]{fig2.ps}
\\
\caption{Spectral energy distribution of $\alpha$ Tau for 1 GHz $\leq \nu \leq$ 1 THz. Our new multi-frequency VLA observations of $\alpha$ Tau (blue circles) were acquired in just two days in February 2011. The red triangles are the previous radio observations of the star which were acquired over many years. The green line is the expected radio emission from the existing hybrid chromosphere and wind model, while the dashed  green line is the expected radio emission from a theoretical Alfv\'en wave driven model atmosphere.}
\label{fig:fig2}
\end{figure*}

@@

In Figure \ref{fig:fig2} we plot the expected radio spectrum of $\alpha$ Tau based on the semi-empirical 1-D chromospheric model of \cite{1999MNRAS.302...37M} embedded in the 1-D wind model of \cite{1998ApJ...503..396R}. The McMurry model, which contains  the photospheric model of \cite{1973ApJ...180...81J}, is based on high S/N Hubble UV spectra and reaches a maximum temperature of 10$^{5}$ K at 1.2 $R_{\star}$. Beyond 1.2 $R_{\star}$ we use Robinson's wind characteristics to describe the outflow velocity where the wind reaches $\sim$80\% of its terminal value by 3 $R_{\star}$. We assume the wind to have a constant temperature of 8,000 K and have a constant ionization fraction, $x_{e}$, of 0.01 throughout. The combination of both atmospheric models result in a predicted excess of radio flux density at high frequencies (i.e., $\nu >$ 30 GHz) but does well in reproducing the VLA flux densities below 30 GHz. The VLA, Institut de Radioastronomie Millim\'{e}trique (IRAM) 30 m-telescope and Berkeley Illinois Maryland Association (BIMA) continuum flux densities confirm that this model predicts a flux excess at even higher frequencies. One possible explanation for this is that the inner atmosphere contains extensive amounts of cooler gas than that predicted by the 1-D static chromospheric model of McMurry. This scenario agrees with the findings of \cite{1994ApJ...423..806W} who conclude that cool regions exist close to the stellar surface with large filling factors i.e., a thermally bifurcated CO-mosphere \citep{1996IAUS..176..371A}.

We also include the predicted radio spectrum from the theoretical Alfv\'{e}n wave-driven outflow model for $\alpha$ Tau \citep{1989AcA....39..51K} in Figure \ref{fig:fig2} to demonstrate how radio observations can test the robustness of theoretical models. This model is based on a fully-ionized outflow at a mass loss rate of  6.3$\times$10$^{-9}$ $M_{\odot}$ yr$^{-1}$, more than two orders of magnitude higher than the more recent estimate given in Table \ref{tab:tab1}. As the radio opacity is proportional to $n _{\rm{e}} n _{\rm{ion}}$, where $n_{\rm{e}}$ and $n_{\rm{ion}}$ are the electron and ion number densities respectively, this model greatly overestimates the actual radio flux density at all VLA wavelengths. The Alfv\'{e}n wave models for $\alpha$ Boo \citep{1988AcA....38..107K} also assume full ionization, and have higher mass loss rates and outflow velocities than the established values given in Table \ref{tab:tab1}, predicting higher flux densities than observed.
